### 2023-09-27 challenges/challenges.tex
Split challenges.tex into multiple files.

\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
                         {\endgroup]\end{quote}}
\long\def\ignore#1{}

\input{../challenges/preliminaries.tex}
\input{../challenges/linearity.tex}
\input{../challenges/scope-extrusion.tex}
\input{../challenges/coinduction.tex}