### 2023-10-17 proof-appendix/proofs.tex

\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
                         {\endgroup]\end{quote}}
\long\def\ignore#1{}

This appendix contains paper proofs (or in some cases, proof sketches), of the lemmas and theorems mentioned in the challenges.
The proofs are all standard, and are in most cases simplified versions of similar proofs in~\cite{picalcbook} and~\cite{Vasconcelos2012}.
These standard references may also be consulted for introductory explanation and examples of the notions discussed in the challenges and for further examples of proofs employing the proof techniques we are interested in.

\input{../proof-appendix/linearity.tex}
\input{../proof-appendix/scope-extrusion.tex}
\input{../proof-appendix/coinduction.tex}