### 2024-01-10 existing-mechanisations/existing-mechanisations.tex
Add table and BibTeX for existing mechanisations.

\subsection{Unprocessed mechanisations}
\begin{itemize}
\item \url{https://doi.org/10.4230/LIPIcs.ITP.2023.28}
\item \url{https://doi.org/10.1007/978-3-030-85315-0_8}
\item \url{https://doi.org/10.4230/LIPIcs.ITP.2022.27}
\item \url{https://doi.org/10.1007/978-3-662-54434-1_34}
\item \url{https://doi.org/10.1145/3354166.3354184}
\item \url{https://doi.org/10.1017/S0960129514000231}
\item \url{https://doi.org/10.1016/j.entcs.2007.11.010}
\item \url{https://doi.org/10.1016/S1571-0661(04)80507-3}
\item \url{https://doi.org/10.1007/BFb0028392}
\item T. F. Melham. 1994. A mechanized theory of the $\pi$-calculus in HOL. Nordic J. of Computing 1, 1 (Spring 1994), 50–76. (\url{https://dl.acm.org/doi/10.5555/640186.640190}, but paper is missing)
\item Otmane Ait-Mohamed. Vérification de l'équivalence du $\pi$-calcul dans HOL. RR-2412, INRIA. 1994 (\url{https://hal.science/inria-00074263/}, in French)
\item \url{https://doi.org/10.1017/S0956796802004653}
\item \url{https://doi.org/10.1007/3-540-45315-6_24}
\item \url{https://inria.hal.science/inria-00072970}
\item \url{https://www.lix.polytechnique.fr/Labo/Dale.Miller/lProlog/examples/pi-calculus/toc.html}
\item \url{https://doi.org/10.1007/BFb0105404}
\item \url{https://doi.org/10.1007/3-540-44755-5_16}
\item \url{https://doi.org/10.1007/3-540-44929-9_30}
\item \url{https://doi.org/10.1007/10721959_33}
\item \url{https://doi.org/10.1016/S0304-3975(00)00095-5}
\item \url{https://doi.org/10.1017/S096012951700010X}
\item \url{https://doi.org/10.1016/j.entcs.2007.11.013}
\item \url{https://doi.org/10.1145/1656242.1656248}
\item \url{https://doi.org/10.1184/R1/6587294.v1}
\item \url{https://doi.org/10.6092/issn.1972-5787/4650}
\item \url{https://doi.org/10.2168/LMCS-5(2:16)2009}
\item \url{https://doi.org/10.4204/EPTCS.203.1}
\item \url{https://doi.org/10.1007/978-3-030-45237-7_17}
\item Uma Zalakain. 2019. Type-checking session-typed $\pi$-calculus with Coq. Master's thesis, University of Glasgow. \url{https://umazalakain.info/static/Z19/msc-thesis.pdf}
\item Adam Michael Petz. 2016. A Semantics for Attestation Protocols using Session Types in Coq. Master's thesis, University of Kansas. \url{http://hdl.handle.net/1808/24137}
\item \url{https://pure.itu.dk/en/publications/representing-session-types}
\item \url{https://doi.org/10.48550/arXiv.1603.03727}
\item \url{https://doi.org/10.1145/3498684}
\item \url{https://doi.org/10.4230/LIPIcs.ITP.2021.15}
\item \url{https://doi.org/10.1007/978-3-319-22102-1_19}
\item \url{https://doi.org/10.1017/S0960129513000170}
\item \url{https://doi.org/10.1007/s10817-015-9336-2}
\item \url{https://doi.org/10.1007/978-3-540-69407-6_33}
\item \url{https://doi.org/10.1016/j.entcs.2007.08.017}
\item \url{https://doi.org/10.1145/3453483.3454041}
\item \url{https://doi.org/10.4204/EPTCS.314.3}
\item \url{https://doi.org/10.7494/csci.2017.18.3.1413}
\item \url{https://doi.org/10.1145/3371074}
\item \url{https://doi.org/10.1145/3158116}
\end{itemize}

A total of 44 papers that we currently know about.

\subsection{Approaches}
\begin{tabular}{l|c|c|c|c|l}
  Reference          & Proof assistant & Linearity & Binders & Coinduction & Limitations \\\hline
  \cite{Tirore:2023} &&&&& \\
  \cite{Cruz-Filipe2021} &&&&& \\
  \cite{Pohjola2022} &&&&& \\
  \cite{Tassarotti2017} &&&&& \\
  \cite{Thiemann2019} &&&&& \\
  \cite{Goto2016} &&&&& \\
  \cite{Affeldt2008} &&&&& \\
  \cite{Scagnetto2002} &&&&& \\
  \cite{Hirschkoff1997} &&&&& \\
  \cite{Melham1994} &&&&& \\
  \cite{Ait-Mohamed1994} &&&&& \\
  \cite{Rockl2003} &&&&& \\
  \cite{Rockl2001} &&&&& \\
  \cite{Henry-Greard1999} &&&&& \\
  \cite{Miller2023} &&&&& \\
  \cite{Gordon1996} &&&&& \\
  \cite{Gay2001} &&&&& \\
  \cite{Despeyroux2000} &&&&& \\
  \cite{Gillard2000} &&&&& \\
  \cite{Honsell2001} &&&&& \\
  \cite{Perera2018} &&&&& \\
  \cite{Watkins2008} &&&&& \\
  \cite{Tiu2010} &&&&& \\
  \cite{Cervesato2007} &&&&& \\
  \cite{Baelde2014} &&&&& \\
  \cite{Bengtson2009} &&&&& \\
  \cite{Orchard2016} &&&&& \\
  \cite{Castro2020} &&&&& \\
  \cite{Zalakain2019} &&&&& \\
  \cite{Petz2016} &&&&& \\
  \cite{Bock2016} &&&&& \\
  \cite{Xi2016} &&&&& \\
  \cite{Hirsch2022} &&&&& \\
  \cite{Cruz-Filipe2021b} &&&&& \\
  \cite{Maksimovic2015} &&&&& \\
  \cite{Parrow2014} &&&&& \\
  \cite{Bengtson2016} &&&&& \\
  \cite{Kahsai2008} &&&&& \\
  \cite{Bengtson2007} &&&&& \\
  \cite{Castro-Perez2021} &&&&& \\
  \cite{Gay2020} &&&&& \\
  \cite{Brady2017} &&&&& \\
  \cite{Hinrichsen2019} &&&&& \\
  \cite{Sergey2017} &&&&&
\end{tabular}

\subsection{Existing experiences and suggestions}
In this section, we collect quotes about experiences attempting to mechanise process calculi and related systems, as well as suggestions proposed by other authors.

\subsubsection{Quotes about challenges encountered}
This section collects quotes about challenges encountered using various techniques for mechanisation.

\subsubsection{Quotes about effort on technical lemmas vs.\ effort on actual content}
This section collects quotes about the amount of effort spent on technical lemmas required by the mechanisation technique versus the amount of effort spent on the theorems of interest.

\subsubsection{Quotes about suggestions for improvements in proof assistant technology}
This section collects quotes suggesting improvements in proof assistant technology and/or mechanisation techniques.