### 2024-01-25 existing-mechanisations/existing-mechanisations.tex
Add references to links in list of unprocessed mechanisations.

\subsection{Unprocessed mechanisations}
\begin{description}
\item[\cite{Tirore:2023}]                 \url{https://doi.org/10.4230/LIPIcs.ITP.2023.28}
\item[\cite{Cruz-Filipe2021}]             \url{https://doi.org/10.1007/978-3-030-85315-0_8}
\item[\cite{Pohjola2022}]                 \url{https://doi.org/10.4230/LIPIcs.ITP.2022.27}
\item[\cite{Tassarotti2017}]              \url{https://doi.org/10.1007/978-3-662-54434-1_34}
\item[\cite{Thiemann2019}]                \url{https://doi.org/10.1145/3354166.3354184}
\item[\cite{Goto2016}]                    \url{https://doi.org/10.1017/S0960129514000231}
\item[\cite{Affeldt2008}]                 \url{https://doi.org/10.1016/j.entcs.2007.11.010}
\item[\cite{Scagnetto2002}]               \url{https://doi.org/10.1016/S1571-0661(04)80507-3}
\item[\cite{Hirschkoff1997}]              \url{https://doi.org/10.1007/BFb0028392}
\item[\cite{Melham1994}]                  T. F. Melham. 1994. A mechanized theory of the $\pi$-calculus in HOL. Nordic J. of Computing 1, 1 (Spring 1994), 50–76. (\url{https://dl.acm.org/doi/10.5555/640186.640190}, but paper is missing)
\item[\cite{Ait-Mohamed1994}]             Otmane Ait-Mohamed. Vérification de l'équivalence du $\pi$-calcul dans HOL. RR-2412, INRIA. 1994 (\url{https://hal.science/inria-00074263/}, in French)
\item[\cite{Rockl2003}]                   \url{https://doi.org/10.1017/S0956796802004653}
\item[\cite{Rockl2001}]                   \url{https://doi.org/10.1007/3-540-45315-6_24}
\item[\cite{Henry-Greard1999}]            \url{https://inria.hal.science/inria-00072970}
\item[\cite{Miller2023}]                  \url{https://www.lix.polytechnique.fr/Labo/Dale.Miller/lProlog/examples/pi-calculus/toc.html}
\item[\cite{Gordon1996}]                  \url{https://doi.org/10.1007/BFb0105404}
\item[\cite{Gay2001}]                     \url{https://doi.org/10.1007/3-540-44755-5_16}
\item[\cite{Despeyroux2000}]              \url{https://doi.org/10.1007/3-540-44929-9_30}
\item[\cite{Gillard2000}]                 \url{https://doi.org/10.1007/10721959_33}
\item[\cite{Honsell2001}]                 \url{https://doi.org/10.1016/S0304-3975(00)00095-5}
\item[\cite{Perera2018}]                  \url{https://doi.org/10.1017/S096012951700010X}
\item[\cite{Watkins2008}]                 \url{https://doi.org/10.1016/j.entcs.2007.11.013}
\item[\cite{Tiu2010}]                     \url{https://doi.org/10.1145/1656242.1656248}
\item[\cite{Cervesato2007}]               \url{https://doi.org/10.1184/R1/6587294.v1}
\item[\cite{Baelde2014}]                  \url{https://doi.org/10.6092/issn.1972-5787/4650}
\item[\cite{Bengtson2009}]                \url{https://doi.org/10.2168/LMCS-5(2:16)2009}
\item[\cite{Orchard2016}]                 \url{https://doi.org/10.4204/EPTCS.203.1}
\item[\cite{Castro2020}]                  \url{https://doi.org/10.1007/978-3-030-45237-7_17}
\item[\cite{Zalakain2019}]                Uma Zalakain. 2019. Type-checking session-typed $\pi$-calculus with Coq. Master's thesis, University of Glasgow. \url{https://umazalakain.info/static/Z19/msc-thesis.pdf}
\item[\cite{Petz2016}]                    Adam Michael Petz. 2016. A Semantics for Attestation Protocols using Session Types in Coq. Master's thesis, University of Kansas. \url{http://hdl.handle.net/1808/24137}
\item[\cite{Bock2016}]                    \url{https://pure.itu.dk/en/publications/representing-session-types}
\item[\cite{Xi2016}]                      \url{https://doi.org/10.48550/arXiv.1603.03727}
\item[\cite{Hirsch2022}]                  \url{https://doi.org/10.1145/3498684}
\item[\cite{Cruz-Filipe2021b}]            \url{https://doi.org/10.4230/LIPIcs.ITP.2021.15}
\item[\cite{Maksimovic2015}]              \url{https://doi.org/10.1007/978-3-319-22102-1_19}
\item[\cite{Parrow2014}]                  \url{https://doi.org/10.1017/S0960129513000170}
\item[\cite{Bengtson2016}]                \url{https://doi.org/10.1007/s10817-015-9336-2}
\item[\cite{Kahsai2008}]                  \url{https://doi.org/10.1007/978-3-540-69407-6_33}
\item[\cite{Bengtson2007}]                \url{https://doi.org/10.1016/j.entcs.2007.08.017}
\item[\cite{Castro-Perez2021}]            \url{https://doi.org/10.1145/3453483.3454041}
\item[\cite{Gay2020}]                     \url{https://doi.org/10.4204/EPTCS.314.3}
\item[\cite{Brady2017}]                   \url{https://doi.org/10.7494/csci.2017.18.3.1413}
\item[\cite{Hinrichsen2019}]              \url{https://doi.org/10.1145/3371074}
\item[\cite{Sergey2017}]                  \url{https://doi.org/10.1145/3158116}
\item[\cite{DBLP:conf/forte/ZalakainD21}] \url{https://doi.org/10.1007/978-3-030-78089-0_9}
\end{description}

A total of 45 papers that we currently know about.

\subsection{Papers written by ``us''}
\subsubsection{Marco and Dawit}
\cite{Tirore:2023} \url{https://doi.org/10.4230/LIPIcs.ITP.2023.28}

\subsubsection{Nobuko}
\cite{Orchard2016} \url{https://doi.org/10.4204/EPTCS.203.1}

\subsubsection{David, Francisco and Nobuko}
\cite{Castro2020} \url{https://doi.org/10.1007/978-3-030-45237-7_17}

\paragraph{+ Lorenzo}
\cite{Castro-Perez2021} \url{https://doi.org/10.1145/3453483.3454041}

\subsection{Approaches}
\begin{tabular}{l|c|c|c|c|l}
  Reference                          & Proof assistant        & Linearity & Binders & Coinduction & Limitations \\\hline
  \cite{Tirore:2023}                 & Coq                    &           &         &             &             \\
  \cite{Cruz-Filipe2021}             & Coq                    &           &         &             &             \\
  \cite{Pohjola2022}                 & HOL4                   &           &         &             &             \\
  \cite{Tassarotti2017}              & Coq                    &           &         &             &             \\
  \cite{Thiemann2019}                & Agda                   &           &         &             &             \\
  \cite{Goto2016}                    & Coq                    &           &         &             &             \\
  \cite{Affeldt2008}                 & Coq                    &           &         &             &             \\
  \cite{Scagnetto2002}               & Coq                    &           &         &             &             \\
  \cite{Hirschkoff1997}              & Coq                    &           &         &             &             \\
  \cite{Melham1994}                  & HOL                    &           &         &             &             \\
  \cite{Ait-Mohamed1994}             & HOL                    &           &         &             &             \\
  \cite{Rockl2003}                   & Isabelle               &           &         &             &             \\
  \cite{Rockl2001}                   & Isabelle               &           &         &             &             \\
  \cite{Henry-Greard1999}            & Coq                    &           &         &             &             \\
  \cite{Miller2023}                  & Abella/$\lambda$Prolog &           &         &             &             \\
  \cite{Gordon1996}                  & HOL                    &           &         &             &             \\
  \cite{Gay2001}                     & Isabelle/HOL           &           &         &             &             \\
  \cite{Despeyroux2000}              & Coq                    &           &         &             &             \\
  \cite{Gillard2000}                 & Coq                    &           &         &             &             \\
  \cite{Honsell2001}                 & Coq                    &           &         &             &             \\
  \cite{Perera2018}                  & Agda                   &           &         &             &             \\
  \cite{Watkins2008}                 & CLF                    &           &         &             &             \\
  \cite{Tiu2010}                     & Bedwyr?                &           &         &             &             \\
  \cite{Cervesato2007}               & CLF                    &           &         &             &             \\
  \cite{Baelde2014}                  & Abella                 &           &         &             &             \\
  \cite{Bengtson2009}                & Isabelle               &           & Nominal &             &             \\
  \cite{Orchard2016}                 & Agda                   &           &         &             &             \\
  \cite{Castro2020}                  & Coq                    &           &         &             &             \\
  \cite{Zalakain2019}                & Coq                    &           &         &             &             \\
  \cite{Petz2016}                    & Coq                    &           &         &             &             \\
  \cite{Bock2016}                    & CLF                    &           &         &             &             \\
  \cite{Xi2016}                      & ATS                    &           &         &             &             \\
  \cite{Hirsch2022}                  & Coq                    &           &         &             &             \\
  \cite{Cruz-Filipe2021b}            & Coq                    &           &         &             &             \\
  \cite{Maksimovic2015}              & Coq                    &           &         &             &             \\
  \cite{Parrow2014}                  & Isabelle               &           & Nominal &             &             \\
  \cite{Bengtson2016}                & Isabelle               &           & Nominal &             &             \\
  \cite{Kahsai2008}                  & Isabelle               &           & Nominal &             &             \\
  \cite{Bengtson2007}                & Isabelle               &           & Nominal &             &             \\
  \cite{Castro-Perez2021}            & Coq                    &           &         &             &             \\
  \cite{Gay2020}                     & Agda                   &           &         &             &             \\
  \cite{Brady2017}                   & Idris                  &           &         &             &             \\
  \cite{Hinrichsen2019}              & Coq                    &           &         &             &             \\
  \cite{Sergey2017}                  & Coq                    &           &         &             &             \\
  \cite{DBLP:conf/forte/ZalakainD21} & Agda                   &           &         &             & 
\end{tabular}

\subsection{Existing experiences and suggestions}
In this section, we collect quotes about experiences attempting to mechanise process calculi and related systems, as well as suggestions proposed by other authors.

\subsubsection{Quotes about challenges encountered}
This section collects quotes about challenges encountered using various techniques for mechanisation.

\subsubsection{Quotes about effort on technical lemmas vs.\ effort on actual content}
This section collects quotes about the amount of effort spent on technical lemmas required by the mechanisation technique versus the amount of effort spent on the theorems of interest.

\subsubsection{Quotes about suggestions for improvements in proof assistant technology}
This section collects quotes suggesting improvements in proof assistant technology and/or mechanisation techniques.