### 2024-01-30 short-paper/lin.tex
first draft of related work abut linearity

\alb{Just a rough attempt to put some words on paper}

In mechanized meta-theory addressing linearity means choosing an
appropriate representation of a linear context. The crux of the matter
is \emph{context splitting} as showcased by the introduction rule of
the tensor. While a linear context can be seen as a multi-set, most
proof assistants have a better support for list processing.  Although
this representation is intuitive, it is overly concrete and requires
establishing dozens and dozens of intermediate lemmas that are foreign
to the mathematics of the problem, say proving type preservation. In
this space, several designs are possible: one can label occurrences of
resources constraining their usage, e.g.~\cite{CicconeP20}, or
\alb{something: ask Franscisco} as in \cite{Castro2020}. Some authors
instead impose a multiset structure over lists,
see~\cite{ChaudhuriLR19,Danielsson12}: still, the effort in developing
the infrastructure is significant. A different and interesting
approach is to bypass the problem of context splitting by adopting
% to  meta-theoretic properties
familiar ideas from linear algorithmic  type checking;
this in known as typing with \emph{leftovers}, as
exemplified in~\cite{DBLP:conf/forte/ZalakainD21}. Whatever the choice,
list-based encodings can be refined to be intrinsically-typed, if the proof assistant supports dependent types, see \cite{Thiemann2019,CicconeP20,RouvoetPKV20}.

For something completely different, one could adopt a
\emph{sub-structural} meta-logical frameworks: here it is the framework
itself that handles resource distribution and the user just has to map
her linear operations to the one offered by the framework. In fact,
the linear function space will handle all context operations
implicitly, including splitting and substitution. One such a framework
is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08}, see the encoding of
session types in~\cite{Bock2016}. Unfortunately, \emph{Clef} does not
yet support meta-theory. A compromise is the so called
\emph{two-level} approach, where one encodes a sub-structural
specification logic in a mainstream proof assistant and then use that
logic to state and prove linear properties, see for a recent
example~\cite{Felty:MSCS21}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "short-paper"
%%% End: