### 2024-01-31 short-paper/lin.tex
small changes to linearity section

\alb{Just a rough attempt to put some words on paper}

In mechanized meta-theory addressing linearity means choosing an
appropriate representation of the linear context. The crux of the matter
is \emph{context splitting} as showcased by the typing rule for parallel composition.
% the introduction rule of the tensor.
While a linear context is best seen as a multi-set, most proof
assistants have better support for list processing.  Although the
latter representation is intuitive, it is overly concrete and may
require establishing dozens and dozens of technical lemmas that are
foreign to the mathematics of the problem under study, say proving
type preservation for session types. In this space, several designs
are possible: one can label occurrences of resources constraining
their usage, e.g.~\cite{CicconeP20}, or \alb{something --- ask
  Francisco} as in \cite{Castro2020}. Some authors instead impose a
multiset structure over lists, see~\cite{ChaudhuriLR19,Danielsson12}:
still, the effort in developing the infrastructure is significant.

A different approach is to bypass the problem of context splitting by
adopting
% to  meta-theoretic properties
familiar ideas from algorithmic linear type checking: this is known as
 \emph{typing with leftovers}, as exemplified
in~\cite{DBLP:conf/forte/ZalakainD21}. Whatever the choice, list-based
encodings can be refined to be intrinsically-typed, if the proof
assistant does support dependent types, see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}.

For something completely different, one could adopt a
\emph{sub-structural} meta-logical framework: here it is the framework
itself that handles resource distribution and the users just have to map
their linear operations to the ones offered by the framework. In fact,
the linear function space will handle all context operations
implicitly, including splitting and substitution. One such a framework
is \emph{Celf}~\cite{Schack-Nielsen:IJCAR08}, see the encoding of
session types in~\cite{Bock2016}. Unfortunately, \emph{Celf} does not
yet fully support the verification of meta-theoretic properties. A compromise is the so called
\emph{two-level} approach, where one encodes a sub-structural
specification logic in a mainstream proof assistant and then use that
logic to state and prove linear properties, see for a recent
example~\cite{Felty:MSCS21}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "short-paper"
%%% End: